---
title: "Assignment 4"
author: "Sungjoo Cho"
date: "`r Sys.Date()`"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## GitHub Link
https://github.com/scho7/SURV727-Assignment4.git

## Packages
```{r message=FALSE, warning=FALSE}
library(bigrquery)
library(DBI)
library(dbplyr)
library(dplyr)
```


## Project
```{r}
# create project object
project <- "surv727-project"

# connect to a public database (Chicago crime database)
con <- dbConnect(
bigrquery::bigquery(),
project = "bigquery-public-data",
dataset = "chicago_crime",
billing = project
)

con
```

```{r}
# check available tables
dbListTables(con)
```

```{sql connection = con}
# crime data in 2016
SELECT *
FROM crime
LIMIT 10;
```



### Write a first query that counts the number of rows of the ‘crime‘ table in the year 2016. Use code chunks with {sql connection = con} in order to write SQL code within the document.

The number of rows of the 'crime' table in the year 2016 is 269841.

```{sql connection = con}
SELECT count(*)
FROM crime
WHERE year = 2016 
LIMIT 10;
```


### Next, count the number of arrests grouped by primary_type in 2016. Note that is a somewhat similar task as above, with some adjustments on which rows should be considered. Sort the results, i.e. list the number of arrests in a descending order.

When sorting the count of the number of arrests grouped by primary_type in 2016 in a descending order, the primary_type NARCOTICS has the highest number of arrests in 2016 of 13327, followed by BATTERY (10332) and THEFT (6522).

```{sql connection = con}
# way 1
SELECT primary_type, count(arrest)
FROM crime
WHERE year = 2016 AND arrest = TRUE
GROUP BY primary_type
ORDER BY COUNT(arrest) DESC;
```

```{r}
# way 2
arr_prm_type_2016 <- "SELECT primary_type, count(arrest)
FROM crime
WHERE year = 2016 AND arrest = TRUE
GROUP BY primary_type
ORDER BY COUNT(arrest) DESC"

dbGetQuery(con, arr_prm_type_2016)
```


### We can also use the date for grouping. Count the number of arrests grouped by hour of the day in 2016. You can extract the latter information from date via EXTRACT(HOUR FROM date). Which time of the day is associated with the most arrests?

When counting the number of arrests grouped by hour of the day in 2016, we can see that time 10 is associated with the most arrests number of 5306, followed by time 11 with the arrests of 5200.

```{sql connection = con}
SELECT EXTRACT(HOUR FROM date), count(*)
FROM crime
WHERE year = 2016 AND arrest = TRUE
GROUP BY EXTRACT(HOUR FROM date)
ORDER BY count(*) DESC;
```


### Focus only on HOMICIDE and count the number of arrests for this incident type, grouped by year. List the results in descending order.

When counting the number of arrests for 'HOMICIDE' incident type, year 2001 has the most highest arrest cases of 430, followed by 2002 with 423 cases.

```{sql connection = con}
SELECT year, count(arrest)
FROM crime
WHERE primary_type = 'HOMICIDE' AND arrest = TRUE
GROUP BY year
ORDER BY count(arrest) DESC;
```


### Find out which districts have the highest numbers of arrests in 2015 and 2016. That is, count the number of arrests in 2015 and 2016, grouped by year and district. List the results in descending order.

In 2016, the district 11 has the highest numbers of arrests with 6575 cases. In 2015, the district 11 also has the highes numbers of arrests with 8974 cases.

```{sql connection = con}
SELECT year, district, count(arrest)
FROM crime
WHERE (year = 2015 OR year = 2016) AND arrest = TRUE
GROUP BY year, district
ORDER BY year DESC, count(arrest) DESC;
```


### Lets switch to writing queries from within R via the DBI package. Create a query object that counts the number of arrests grouped by primary_type of district 11 in year 2016. The results should be displayed in descending order. Execute the query.

```{r}
arr_dis11_2016 <- "SELECT primary_type, count(arrest)
FROM crime
WHERE (district = 11 AND year = 2016) AND arrest = TRUE
GROUP BY primary_type
ORDER BY count(arrest) DESC"

dbGetQuery(con, arr_dis11_2016)
```


### Try to write the very same query, now using the dbplyr package. For this, you need to first map the crime table to a tibble object in R.

```{r}
crime <- tbl(con, "crime")
str(crime)
class(crime)
```


### Again, count the number of arrests grouped by primary_type of district 11 in year 2016, now using dplyr syntax.

```{r}
crime %>%
  filter(district == 11, year == 2016, arrest == TRUE) %>%
  group_by(primary_type) %>%
  summarize(count = n()) %>% 
  arrange(desc(count))
```


### Count the number of arrests grouped by primary_type and year, still only for district 11. Arrange the result by year.

### Assign the results of the query above to a local R object.

```{r}
query <- crime %>%
  filter(district == 11, arrest == TRUE) %>%
  group_by(year, primary_type) %>%
  summarize(count = n()) %>% 
  arrange(desc(year))

query

# translate it into SQL
show_query(query)
```


```{r}

```


### Confirm that you pulled the data to the local environment by displaying the first ten rows of the saved data set.

```{r}
sql_data <- 
  crime %>%
  head(10) %>%
  collect()

str(sql_data)
```


### Close the connection.

```{r}
dbDisconnect(con)
```




